---
title: "ESM 260 - Problem Set 1"
author: "Laura Ingulsrud"
date: "1/26/2020"
output: 
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 5, fig.width = 8)
```

```{r packages, include=FALSE}

# Load necessary packages
library(tidyverse)
library(sanzo)

```

```{r data, include=FALSE}

# Read in data
density <- read_csv("density_adultfish.csv")
density_other <- read_csv("density_other.csv")
recruited <- read_csv("recruited_fish.csv")

```

####1. Mean Density of Adult Fish

```{r #1, include=FALSE}

# Create bar charts for each organism that displays the mean density of the organism (y-axis) at Rarotonga and Aitutaki for 2009 and 2018. Add the 95% confidence intervals to each graph.

################# Surgeonfish #################

surgeonfish <- density %>% 
  filter(organism == "Surgeonfish")

surgeonfish_plot <- ggplot(surgeonfish, aes(x = location, y = density, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density - CI, ymax = density + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Adult Surgeonfish") +
  scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c074"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeonfish_plot  

################# Coral Trout #################

coraltrout <- density %>% 
  filter(organism == "Coral Trout")

coraltrout_plot <- ggplot(coraltrout, aes(x = location, y = density, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density - CI, ymax = density + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Adult Coral Trout") +
  scale_y_continuous(expand = c(0,0), limits = c(-20,200), breaks = seq(0,200,50)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c074"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

coraltrout_plot

################# Spotted Damselfish #################

spotteddamselfish <- density %>% 
  filter(organism == "Spotted Damselfish")

spotteddamselfish_plot <- ggplot(spotteddamselfish, aes(x = location, y = density, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density - CI, ymax = density + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Adult Spotted Damselfish") +
  scale_y_continuous(expand = c(0,0), limits = c(-40,1000), breaks = seq(0,1000,200)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c074"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

spotteddamselfish_plot

################# Yellow Damselfish #################

yellowdamselfish <- density %>% 
  filter(organism == "Yellow Damselfish")

yellowdamselfish_plot <- ggplot(yellowdamselfish, aes(x = location, y = density, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density - CI, ymax = density + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Adult Yellow Damselfish") +
  scale_y_continuous(expand = c(0,0), limits = c(0,110), breaks = seq(0,100,20)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c074"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

yellowdamselfish_plot

################# Mounding Corals #################

moundingcoral <- density_other %>% 
  filter(organism == "Mounding Corals")

moundingcoral_plot <- ggplot(moundingcoral, aes(x = location, y = density_habitat, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_habitat - CI, ymax = density_habitat + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Mounding Corals") +
  scale_y_continuous(expand = c(0,0), limits = c(0,25), breaks = seq(0,25,5)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c102"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

moundingcoral_plot

################# Coarse-branching Corals #################

coarsecoral <- density_other %>% 
  filter(organism == "Coarse-branching Corals")

coarsecoral_plot <- ggplot(coarsecoral, aes(x = location, y = density_habitat, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_habitat - CI, ymax = density_habitat + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Coarse-branching Corals") +
  scale_y_continuous(expand = c(0,0), limits = c(0,4), breaks = seq(0,4,1)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c102"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

coarsecoral_plot

################# Fine-branching Corals #################

finecoral <- density_other %>% 
  filter(organism == "Fine-branching Corals")

finecoral_plot <- ggplot(finecoral, aes(x = location, y = density_habitat, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_habitat - CI, ymax = density_habitat + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Fine-branching Corals") +
  scale_y_continuous(expand = c(0,0), limits = c(-0.7,4.3), breaks = seq(0,4,1)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c102"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

finecoral_plot

################# Sea Anemones #################

anemone <- density_other %>% 
  filter(organism == "Sea Anemones")

anemone_plot <- ggplot(anemone, aes(x = location, y = density_habitat, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_habitat - CI, ymax = density_habitat + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Sea Anemones") +
  scale_y_continuous(expand = c(0,0), limits = c(0,2), breaks = seq(0,2,0.5)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c102"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

anemone_plot

################# Algae #################

algae <- density_other %>% 
  filter(organism == "Algae")

algae_plot <- ggplot(algae, aes(x = location, y = density_habitat, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = density_habitat - CI, ymax = density_habitat + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Organisms~(per~100~m^2)),
    title = "Mean Density of Algae") +
  scale_y_continuous(expand = c(0,0), limits = c(-0.5,6.5), breaks = seq(0,6,2)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c102"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

algae_plot

```

```{r, echo=FALSE}

surgeonfish_plot 

```

There were significantly less adult Surgeonfish in Rarotonga compared to Aitutaki in both 2009 and 2018, although this difference is more drastic in 2018. In Rarotonga, adult Surgeonfish significantly decreased from 2009 to 2018.

```{r, echo=FALSE}

coraltrout_plot

```

There were significantly less adult Coral Trout in Rarotonga compared to Aitutaki in 2009. In Aitutaki and Rarotonga, adult Coral Trout significantly decreased from 2009 to 2018.

```{r, echo=FALSE}

spotteddamselfish_plot

```

There were significantly less adult Spotted Damselfish in Rarotonga compared to Aitutaki in both 2009 and 2018. In Aitutaki and Rarotonga, adult Spotted Damselfish significantly decreased from 2009 to 2018.

```{r, echo=FALSE}

yellowdamselfish_plot

```

In Aitutaki and Rarotonga, adult Yellow Damselfish significantly increased from 2009 to 2018. 

```{r, echo=FALSE}

moundingcoral_plot

```

There are no statistically significant differences in mounding coral density across lagoons and years.

```{r, echo=FALSE}

coarsecoral_plot

```

In Rarotonga, Coarse-branching Corals significantly decreased from 2009 to 2018. Coarse-branching Corals density in Rarotonga is significantly less than in Aitutaki in 2018.

```{r, echo=FALSE}

finecoral_plot

```

In Rarotonga, Fine-branching Corals significantly decreased from 2009 to 2018. Coarse-branching Corals density in Rarotonga is significantly less than in Aitutaki in 2018.

```{r, echo=FALSE}

anemone_plot

```

In Rarotonga and Aitutaki, Sea Anemones significantly decreased from 2009 to 2018. Sea Anemone densities in Rarotonga significantly less than in Aitutaki in both 2009 and 2018.

```{r, echo=FALSE}

algae_plot

```

In Rarotonga, Algae significantly increased from 2009 to 2018. Algae density in Rarotonga is significantly greater than in Aitutaki in 2018.

####2. Mean Density of Recruited Young Fish

```{r #2, include=FALSE}

# Create bar charts for each recruit (young) of each fish species that displays the mean density of the organism (y-axis) at Rarotonga and Aitutaki for 2009 and 2018. Add the 95% confidence intervals to each graph.

################# Surgeonfish #################

surgeonfish_recruit <- recruited %>% 
  filter(fish == "Surgeonfish") %>% 
  filter(number != 0)

surgeonfish_recruit_plot <- ggplot(surgeonfish_recruit, aes(x = location, y = number, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = number - CI, ymax = number + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Fish~Recruits~(per~1~m^2~of~microhabitat)),
    title = "Mean Density of Recruited Surgeonfish") +
  scale_y_continuous(expand = c(0,0), limits = c(0,45), breaks = seq(0,45,10)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c047"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeonfish_recruit_plot  

################# Coral Trout #################

coraltrout_recruit <- recruited %>% 
  filter(fish == "Coral Trout") %>% 
  filter(number != 0)

coraltrout_recruit_plot <- ggplot(coraltrout_recruit, aes(x = location, y = number, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = number - CI, ymax = number + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Fish~Recruits~(per~1~m^2~of~microhabitat)),
    title = "Mean Density of Recruited Coral Trout") +
  scale_y_continuous(expand = c(0,0), limits = c(-1.6,15), breaks = seq(0,15,5)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c047"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

coraltrout_recruit_plot  

################# Spotted Damselfish #################

spotted_recruit <- recruited %>% 
  filter(fish == "Spotted Damselfish") %>% 
  filter(number != 0)

spotted_recruit_plot <- ggplot(spotted_recruit, aes(x = location, y = number, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = number - CI, ymax = number + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Fish~Recruits~(per~1~m^2~of~microhabitat)),
    title = "Mean Density of Recruited Spotted Damselfish") +
  scale_y_continuous(expand = c(0,0), limits = c(0,21), breaks = seq(0,20,5)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c047"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

spotted_recruit_plot 

################# Yellow Damselfish #################

yellow_recruit <- recruited %>% 
  filter(fish == "Yellow Damselfish") %>% 
  filter(number != 0)

yellow_recruit_plot <- ggplot(yellow_recruit, aes(x = location, y = number, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  geom_errorbar(aes(ymin = number - CI, ymax = number + CI), position = position_dodge(0.9), width = 0.3) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Fish~Recruits~(per~1~m^2~of~microhabitat)),
    title = "Mean Density of Recruited Yellow Damselfish") +
  scale_y_continuous(expand = c(0,0), limits = c(0,52), breaks = seq(0,50,10)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c047"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

yellow_recruit_plot 

```

```{r, echo=FALSE}

surgeonfish_recruit_plot 

```

There are no statistically significant differences in mounding coral density across lagoons and years.

```{r, echo=FALSE}

coraltrout_recruit_plot

```

In Rarotonga, recruited Coral Trout significantly decreased from 2009 to 2018. There were significantly less recruited Coral Trout in Rarotonga compared to Aitutaki in 2018.

```{r, echo=FALSE}

spotted_recruit_plot

```

There are no statistically significant differences in mounding coral density across lagoons and years.

```{r, echo=FALSE}

yellow_recruit_plot

```

There were significantly more recruited Yellow Damselfish in Rarotonga compared to Aitutaki in both 2009 and 2018. In both Aitutaki and Rarotonga, recruited Yellow Damselfish significantly increased from 2009 to 2018.

####3. Density of Fish Recruits per 100m^2 Area of Reef

```{r #3, include=FALSE}

# Multiply number of fish recruits per 1 m2 of microhabitat by the density of "other organisms" (microhabitat) per 100 m2 of reef area to get the density of fish recruits per 100 m2 of reef area

################# Surgeonfish (microhabitat = Fine-branching Coral) #################

surgeonfish_recruit_density <- (surgeonfish_recruit$number * finecoral$density_habitat)

surgeonfish_all <- cbind(surgeonfish_recruit_density, surgeonfish)

surgeonfish_r2 <- surgeonfish_all %>% 
  summarize(R2 = summary(lm(surgeonfish_recruit_density ~ density))$r.squared)
# R2 = 0.96754

# Create XY scatter plot of density of recruits and adults
surgeonfish_all_plot <- ggplot(surgeonfish_all, aes(
                                                    x = surgeonfish_recruit_density, 
                                                    y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Fish~Recruits~(per~100~m^2~of~reef~area)), 
    y = expression(Density~of~Adult~Fish~(per~100~m^2~of~reef~area)),
    title = expression(Density~of~Surgeonfish~Recruits~and~Adults~per~100~m^2~of~reef~area),
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 20, y = 350, label = paste("R^2 ==", format(surgeonfish_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c091")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeonfish_all_plot

################# Coral Trout (microhabitat = Coarse-branching Coral) #################

coraltrout_recruit_density <- (coraltrout_recruit$number * coarsecoral$density_habitat)

coraltrout_all <- cbind(coraltrout_recruit_density, coraltrout)

coraltrout_r2 <- coraltrout_all %>% 
  summarize(R2 = summary(lm(coraltrout_recruit_density ~ density))$r.squared)
# R2 = 0.8957753

# Create XY scatter plot of density of recruits and adults
coraltrout_all_plot <- ggplot(coraltrout_all, aes(
                                                    x = coraltrout_recruit_density, 
                                                    y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Fish~Recruits~(per~100~m^2~of~reef~area)), 
    y = expression(Density~of~Adult~Fish~(per~100~m^2~of~reef~area)),
    title = expression(Density~of~Coral~Trout~Recruits~and~Adults~per~100~m^2~of~reef~area),
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 10, y = 150, label = paste("R^2 ==", format(coraltrout_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c091")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

coraltrout_all_plot

################# Spotted Damselfish (microhabitat = Sea Anemones) #################

spotted_recruit_density <- (spotted_recruit$number * anemone$density_habitat)

spotted_all <- cbind(spotted_recruit_density, spotteddamselfish)

spotted_r2 <- spotted_all %>% 
  summarize(R2 = summary(lm(spotted_recruit_density ~ density))$r.squared)
# R2 = 0.9434432

# Create XY scatter plot of density of recruits and adults
spotted_all_plot <- ggplot(spotted_all, aes(
                                            x = spotted_recruit_density, 
                                            y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Fish~Recruits~(per~100~m^2~of~reef~area)), 
    y = expression(Density~of~Adult~Fish~(per~100~m^2~of~reef~area)),
    title = expression(Density~of~Spotted~Damselfish~Recruits~and~Adults~per~100~m^2~of~reef~area),
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 8, y = 750, label = paste("R^2 ==", format(spotted_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c091")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

spotted_all_plot

################# Yellow Damselfish (microhabitat = Mounding Coral) #################

yellow_recruit_density <- (yellow_recruit$number * moundingcoral$density_habitat)

yellow_all <- cbind(yellow_recruit_density, yellowdamselfish)

yellow_r2 <- yellow_all %>% 
  summarize(R2 = summary(lm(yellow_recruit_density ~ density))$r.squared)
# R2 = 0.401724

# Create XY scatter plot of density of recruits and adults
yellow_all_plot <- ggplot(yellow_all, aes(
                                          x = spotted_recruit_density, 
                                          y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Fish~Recruits~(per~100~m^2~of~reef~area)), 
    y = expression(Density~of~Adult~Fish~(per~100~m^2~of~reef~area)),
    title = expression(Density~of~Yellow~Damselfish~Recruits~and~Adults~per~100~m^2~of~reef~area),
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 16, y = 80, label = paste("R^2 ==", format(yellow_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c091")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

yellow_all_plot


```

```{r, echo=FALSE}

surgeonfish_all_plot

```

Surgeonfish shows a strong positive correlation in adult and recruit densities.

```{r, echo=FALSE}

coraltrout_all_plot

```

Coral Trout shows a strong positive correlation in adult and recruit densities.

```{r, echo=FALSE}

spotted_all_plot

```

Spotted Damselfish shows a strong positive correlation in adult and recruit densities.

```{r, echo=FALSE}

yellow_all_plot

```

Yellow Damselfish shows a weaker negative correlation in adult and recruit densities.

####4. Relationship between density of adults and density of microhabitat used by young

```{r #4, include=FALSE}

# Create XY scatter plot of density of adults and density of microhabitat used by their young

################# Surgeonfish vs fine-branching coral #################

surgeon_fine <- cbind(surgeonfish, finecoral$density_habitat)

surgeon_fine_r2 <- surgeon_fine %>% 
  summarize(R2 = summary(lm(density ~ `finecoral$density_habitat`))$r.squared)
# R2 = 0.9646357

surgeon_fine_plot <- ggplot(surgeon_fine, aes(x = finecoral$density_habitat, y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Young~Fish~Microhabitat~(1~m^2~per~100~m^2~of~reef~area)),
    y = expression(Density~of~Adult~Fish~(Number~per~100~m^2~of~reef)), 
    title = "Density of Adult Surgeonfish and Fine-branching Coral",
    color = "Lagoon",
    shape = "Year") +
  annotate(geom = "text", x = 1, y = 350, label = paste("R^2 ==", format(surgeon_fine_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c070")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeon_fine_plot

################# Coral Trout vs coarse-branching coral #################

coral_coarse <- cbind(coraltrout, coarsecoral$density_habitat)

coral_coarse_r2 <- coral_coarse %>% 
  summarize(R2 = summary(lm(density ~ `coarsecoral$density_habitat`))$r.squared)
# R2 = 0.7761476

coral_coarse_plot <- ggplot(coral_coarse, aes(x = coarsecoral$density_habitat, y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Young~Fish~Microhabitat~(1~m^2~per~100~m^2~of~reef~area)),
    y = expression(Density~of~Adult~Fish~(Number~per~100~m^2~of~reef)), 
    title = "Density of Adult Coral Trout and Coarse-branching Coral",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 1.25, y = 150, label = paste("R^2 ==", format(coral_coarse_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c070")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

coral_coarse_plot

################# Spotted Damselfish vs sea anemones #################

spotted_anemone <- cbind(spotteddamselfish, anemone$density_habitat)

spotted_anemone_r2 <- spotted_anemone %>% 
  summarize(R2 = summary(lm(density ~ `anemone$density_habitat`))$r.squared)
# R2 = 0.9482483

spotted_anemone_plot <- ggplot(spotted_anemone, aes(x = anemone$density_habitat, y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Young~Fish~Microhabitat~(1~m^2~per~100~m^2~of~reef~area)),
    y = expression(Density~of~Adult~Fish~(Number~per~100~m^2~of~reef)), 
    title = "Density of Adult Spotted Damselfish and Sea Anemones",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 0.8, y = 875, label = paste("R^2 ==", format(spotted_anemone_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c070")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

spotted_anemone_plot

################# Yellow Damselfish vs mounding coral #################

yellow_mounding <- cbind(yellowdamselfish, moundingcoral$density_habitat)

yellow_mounding_r2 <- yellow_mounding %>% 
  summarize(R2 = summary(lm(density ~ `moundingcoral$density_habitat`))$r.squared)
# R2 = 0.0001074609

yellow_mounding_plot <- ggplot(yellow_mounding, aes(x = moundingcoral$density_habitat, y = density)) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  labs(
    x = expression(Density~of~Young~Fish~Microhabitat~(1~m^2~per~100~m^2~of~reef~area)),
    y = expression(Density~of~Adult~Fish~(Number~per~100~m^2~of~reef)), 
    title = "Density of Adult Yellow Damselfish and Mounding Coral",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 15, y = 85, label = paste("R^2 ==", format(yellow_mounding_r2, digits = 2)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c070")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

yellow_mounding_plot

```

```{r, echo=FALSE}

surgeon_fine_plot

```

Adult Surgeonfish shows a strong positive correlation between its density and its young's microhabitat (Fine-branching Coral) density. In Aitutaki from 2009 to 2018, this relationship indicates a trend of increasing young fish microhabitat leading to increasing adult fish associated with that microhabitat. In Rarotonga from 2009 to 2018 however, this trend reverses, with both Fine-branching Coral and adult Surgeonfish decreasing over time.

```{r, echo=FALSE}

coral_coarse_plot

```

Adult Coral Trout shows a positive correlation between its density and its young's microhabitat (Coarse-branching Coral) density. In both Aitutaki and Rarotonga from 2009 to 2018, this relationship indicates a trend of both Coarse-branching Coral and adult Coral Trout decreasing over time. 

```{r, echo=FALSE}

spotted_anemone_plot

```

Adult Spotted Damselfish shows a strong positive correlation between its density and its young's microhabitat (Sea Anemones) density. In both Aitutaki and Rarotonga from 2009 to 2018, this relationship indicates a trend of both Sea Anemones and adult Spotted Damselfish decreasing over time.

```{r, echo=FALSE}

yellow_mounding_plot

```

There is no significant correlation between adult Yellow Damselfish density and its young's microhabitat (Mounding Coral) density. This relationship indicates there is no trend associated with Mounding Coral and adult Yellow Damselfish.


####5. Density of adults per area of microhabitat used by young

```{r #5, include=FALSE}

################# Surgeonfish #################

# Find density of adults/microhabitat (fine-branching coral)
surgeon_fine$adult_micro <- surgeon_fine$density/surgeon_fine$`finecoral$density_habitat`

surgeon_micro_plot <- ggplot(surgeon_fine, aes(x = location, y = adult_micro, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Fish~(per~100~m^2~of~Microhabitat)),
    title = "Density of Adult Surgeonfish per Fine-branching Coral Microhabitat Area") +
  scale_y_continuous(expand = c(0,0), limits = c(0,150)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c119"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(size = 12, hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeon_micro_plot  

################# Coral Trout #################

# Find density of adults/microhabitat (coarse-branching coral)
coral_coarse$adult_micro <- coral_coarse$density/coral_coarse$`coarsecoral$density_habitat`

coral_micro_plot <- ggplot(coral_coarse, aes(x = location, y = adult_micro, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Fish~(per~100~m^2~of~Microhabitat)),
    title = "Density of Adult Coral Trout per Coarse-branching Coral Microhabitat Area") +
  scale_y_continuous(expand = c(0,0), limits = c(0,55)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c119"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(size = 12, hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

coral_micro_plot  

################# Spotted Damselfish #################

# Find density of adults/microhabitat (sea anemones)
spotted_anemone$adult_micro <- spotted_anemone$density/spotted_anemone$`anemone$density_habitat`

spotted_micro_plot <- ggplot(spotted_anemone, aes(x = location, y = adult_micro, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Fish~(per~100~m^2~of~Microhabitat)),
    title = "Density of Adult Spotted Damselfish per Sea Anemone Microhabitat Area") +
  scale_y_continuous(expand = c(0,0), limits = c(0,550)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c119"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(size = 12, hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

spotted_micro_plot 

################# Yellow Damselfish #################

# Find density of adults/microhabitat (mounding coral)
yellow_mounding$adult_micro <- yellow_mounding$density/yellow_mounding$`moundingcoral$density_habitat`

yellow_micro_plot <- ggplot(yellow_mounding, aes(x = location, y = adult_micro, fill = location)) +
  geom_bar(stat = "identity", color = "black", position = "dodge", alpha = 0.9) +
  facet_wrap(~year) +
  labs(
    x = "Lagoon", 
    y = expression(Mean~Density~of~Adult~Fish~(per~100~m^2~of~Microhabitat)),
    title = "Density of Adult Yellow Damselfish per Mounding Coral Microhabitat Area") +
  scale_y_continuous(expand = c(0,0), limits = c(0,5.5)) +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_manual(values = (sanzo.duo("c119"))) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(size = 12, hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

yellow_micro_plot 

```

```{r, echo=FALSE}

surgeon_micro_plot

```

Between Rarotonga and Aitutaki, Surgeonfish show apparent differences in the density of adults per area of juvenile microhabitat in both 2009 and 2018, with densities being lower in Rarotonga compared to Aitutaki. Surgeonfish show an apparent decrease in the density of adults per area of juvenile habitat in both lagoons between 2009 and 2018.  

```{r, echo=FALSE}

coral_micro_plot

```

Between Rarotonga and Aitutaki, Coral Trout show apparent differences in the density of adults per area of juvenile microhabitat in both 2009 and 2018, with densities being lower in Rarotonga compared to Aitutaki. Coral Trout show an apparent decrease in the density of adults per area of juvenile habitat in both lagoons between 2009 and 2018, with a large decrease appearing in Rarotonga between 2009 and 2018.  

```{r, echo=FALSE}

spotted_micro_plot

```

Spotted Damselfish show apparent differences in the density of adults per area of juvenile microhabitat in 2018, with densities being much lower in Rarotonga compared to Aitutaki, but densities are pretty similar between the two lagoons in 2009. Spotted Damselfish show an apparent decrease in the density of adults per area of juvenile habitat in both lagoons between 2009 and 2018, with a large decreases appearing in Rarotonga between 2009 and 2018. 

```{r, echo=FALSE}

yellow_micro_plot

```

Yellow Damselfish show an apparent increase in the density of adults per area of juvenile habitat in both lagoons between 2009 and 2018, although there isn't much difference between the two lagoons in both years.

####6. Compare densities of adults

```{r #6, include=FALSE}

################# Surgeonfish vs Coral Trout #################

# Add coral trout density to surgeonfish tibble
surgeon_trout <- cbind(surgeonfish, coraltrout$density)

# Find R2
surgeon_trout_r2 <- surgeon_trout %>% 
  summarize(R2 = summary(lm(density ~ `coraltrout$density`))$r.squared)
  # R2 = 0.3275856

surgeon_trout_plot <- ggplot(surgeon_trout, aes(x = density, y = `coraltrout$density`)) +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  labs(
    x = expression(Adult~Surgeonfish~Density~(Number~per~100~m^2~of~reef)),
    y = expression(Adult~Coral~Trout~Density~(Number~per~100~m^2~of~reef)),
    title = "Surgeonfish and Coral Trout Densities",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 100, y = 150, label = paste("R^2 ==", format(surgeon_trout_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c114")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeon_trout_plot

################# Surgeonfish vs Spotted Damselfish #################

# Add spotted damselfish density to surgeonfish tibble
surgeon_spotted <- cbind(surgeonfish, spotteddamselfish$density)

# Find R2
surgeon_spotted_r2 <- surgeon_spotted %>% 
  summarize(R2 = summary(lm(density ~ `spotteddamselfish$density`))$r.squared)
  # R2 = 0.410623

surgeon_spotted_plot <- ggplot(surgeon_spotted, aes(x = density, y = `spotteddamselfish$density`)) +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  labs(
    x = expression(Adult~Surgeonfish~Density~(Number~per~100~m^2~of~reef)),
    y = expression(Adult~Spotted~Damselfish~Density~(Number~per~100~m^2~of~reef)),
    title = "Surgeonfish and Spotted Damselfish Densities",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 100, y = 750, label = paste("R^2 ==", format(surgeon_spotted_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c114")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeon_spotted_plot

################# Surgeonfish vs Yellow Damselfish #################

# Add yellow damselfish density to surgeonfish tibble
surgeon_yellow <- cbind(surgeonfish, yellowdamselfish$density)

# Find R2
surgeon_yellow_r2 <- surgeon_yellow %>% 
  summarize(R2 = summary(lm(density ~ `yellowdamselfish$density`))$r.squared)
  # R2 = 0.2015835

surgeon_yellow_plot <- ggplot(surgeon_yellow, aes(x = density, y = `yellowdamselfish$density`)) +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  labs(
    x = expression(Adult~Surgeonfish~Density~(Number~per~100~m^2~of~reef)),
    y = expression(Adult~Yellow~Damselfish~Density~(Number~per~100~m^2~of~reef)),
    title = "Surgeonfish and Yellow Damselfish Densities",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 200, y = 85, label = paste("R^2 ==", format(surgeon_yellow_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c114")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

surgeon_yellow_plot

################# Coral Trout vs Spotted Damselfish #################

# Add spotted damselfish density to coral trout tibble
trout_spotted <- cbind(coraltrout, spotteddamselfish$density)

# Find R2
trout_spotted_r2 <- trout_spotted %>% 
  summarize(R2 = summary(lm(density ~ `spotteddamselfish$density`))$r.squared)
  # R2 = 0.9704437

trout_spotted_plot <- ggplot(trout_spotted, aes(x = density, y = `spotteddamselfish$density`)) +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  labs(
    x = expression(Adult~Coral~Trout~Density~(Number~per~100~m^2~of~reef)),
    y = expression(Adult~Spotted~Damselfish~Density~(Number~per~100~m^2~of~reef)),
    title = "Coral Trout and Spotted Damselfish Densities",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 50, y = 750, label = paste("R^2 ==", format(trout_spotted_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c114")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

trout_spotted_plot

################# Coral Trout vs Yellow Damselfish #################

# Add yellow damselfish density to coral trout tibble
trout_yellow <- cbind(coraltrout, yellowdamselfish$density)

# Find R2
trout_yellow_r2 <- trout_yellow %>% 
  summarize(R2 = summary(lm(density ~ `yellowdamselfish$density`))$r.squared)
  # R2 = 0.8914587

trout_yellow_plot <- ggplot(trout_yellow, aes(x = density, y = `yellowdamselfish$density`)) +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  labs(
    x = expression(Adult~Coral~Trout~Density~(Number~per~100~m^2~of~reef)),
    y = expression(Adult~Yellow~Damselfish~Density~(Number~per~100~m^2~of~reef)),
    title = "Coral Trout and Yellow Damselfish Densities",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 125, y = 80, label = paste("R^2 ==", format(trout_yellow_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c114")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

trout_yellow_plot

################# Spotted Damselfish vs Yellow Damselfish #################

# Add yellow damselfish density to spotted damselfish tibble
spotted_yellow <- cbind(spotteddamselfish, yellowdamselfish$density)

# Find R2
spotted_yellow_r2 <- spotted_yellow %>% 
  summarize(R2 = summary(lm(density ~ `yellowdamselfish$density`))$r.squared)
  # R2 = 0.7656169

spotted_yellow_plot <- ggplot(spotted_yellow, aes(x = density, y = `yellowdamselfish$density`)) +
  geom_point(size = 5, aes(color = location, shape = factor(year))) +
  geom_smooth(method = lm, se = F, color = "slategrey") +
  labs(
    x = expression(Adult~Spotted~Damselfish~Density~(Number~per~100~m^2~of~reef)),
    y = expression(Adult~Yellow~Damselfish~Density~(Number~per~100~m^2~of~reef)),
    title = "Spotted Damselfish and Yellow Damselfish Densities",
    color = "Lagoon",
    shape = "Year")+
  annotate(geom = "text", x = 625, y = 80, label = paste("R^2 ==", format(spotted_yellow_r2, digits = 4)), color = "red4", family = "serif", parse = TRUE) +
  scale_color_manual(values = sanzo.duo("c114")) +
  theme_bw() +
  theme(text = element_text(family = "serif")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"))

spotted_yellow_plot

```

```{r, echo=FALSE}

surgeon_trout_plot

```

There is a weak positive correlation between adult Coral Trout and adult Surgeonfish densities overall. In Aitutaki from 2009 to 2018, this relationship indicates a trend of increasing Surgeonfish but decreasing Coral Trout. In Rarotonga from 2009 to 2018, this relationship indicates a negative trend, with both Coral Trout and Surgeonfish decreasing over time.

```{r, echo=FALSE}

surgeon_spotted_plot

```

There is a weak positive correlation between adult Spotted Damselfish and adult Surgeonfish densities overall. In Aitutaki from 2009 to 2018, this relationship indicates a trend of increasing Surgeonfish but decreasing Spotted Damselfish. In Rarotonga from 2009 to 2018, this relationship indicates a negative trend, with both Spotted Damselfish and Surgeonfish decreasing over time.

```{r, echo=FALSE}

surgeon_yellow_plot

```

There is a weak negative correlation between adult Yellow Damselfish and adult Surgeonfish densities overall. In Aitutaki from 2009 to 2018, this relationship indicates a positive trend of increasing Surgeonfish and Yellow Damselfish. In Rarotonga from 2009 to 2018, this trend shows Yellow Damselfish increasing, while Surgeonfish is decreasing over time.

```{r, echo=FALSE}

trout_spotted_plot

```
There is a strong positive correlation between adult Coral Trout and adult Spotted Damselfish densities overall. In both Aitutaki and Rarotonga from 2009 to 2018, this relationship indicates a trend of decreasing Coral Trout and Spotted Damselfish over time.

```{r, echo=FALSE}

trout_yellow_plot

```

There is a strong negative correlation between adult Coral Trout and adult Yellow Damselfish densities overall. In both Aitutaki and Rarotonga from 2009 to 2018, this relationship indicates a trend of decreasing Coral Trout but increasing Yellow Damselfish over time.

```{r, echo=FALSE}

spotted_yellow_plot

```

There is a negative correlation between adult Spotted Damselfish and adult Yellow Damselfish densities overall. In both Aitutaki and Rarotonga from 2009 to 2018, this relationship indicates a trend of decreasing Spotted Damselfish but increasing Yellow Damselfish over time.